Supplementary Information

### 2021-07-28 doc/manuscript/response_to_reviews.Rmd

**Thank you!**

More general comments:

Data:
I am not sure why there is no data shown on any of the figures presented. Inclusion of the raw data should help support the results of the analysis. Since they are omitted, it makes me (as a reader) question if they are intentionally not shown as a result of some discrepancy. At the very least, I would like to see data included in a panel of each figure, and a comment as to why it was omitted in the rest.

**In general, data were not presented because (1) there's inherently a lot of variation in tree-ring (or any tree growth) data, so showing the data makes it tough to visualize the trends; (2) figures are showing partial effects and so showing the raw data would be difficult because in multivariate analyses, any single factor is showing only one part of the variance and so it's hard to meaningfully show the "raw" data, and (3) we're packing so much into all the figures that adding data points would make them too busy. To show some example data, we have modified Figure 1 to show raw data with model fits for one site-species combination (replacing an example showing just model fits for all species at one site). This effectively implements the suggestion to show example data for one panel in each figure, as Figs 3,5,and 6 include the panels in this example figure for all site-species combinations. We note also that raw data are visualized in Figs. 2 and S6-S9, and parallel plots for all sites are available in the GitHub repository (https://github.com/EcoClimLab/ForestGEO-tree-rings) and archived in Zenodo.**

@@

Census:
I think the article mentions that most of these forest plots have repeat census data. How was this data used in the methods? If there were multiple census DBH measurements, how were these all used to constrain DBH time series?

**We have added a short paragraph to Appendix S2 specifying how DBH measurements were used: "DBH measurements were taken at the time of coring or -- for some of the trees cored within ForestGEO plots -- obtained from the most recent ForestGEO census prior to coring. For these trees, we applied the outside-in approach to DBH reconstruction (detailed below). For a minority of trees (n=50 at LDW), DBH measurements were not available, requiring use of inside-out approach to DBH reconstruction (detailed below)." We did not use multiple census measurements to constrain the DBH time series (a worthy endeavor, but beyond the scope of this project).**

### 2021-07-28 doc/manuscript/response_to_reviews.Rmd

*increased mortality--> decreased competition --> increased growth. This should be mentioned in the discussion.*

P19-20 (especially L301-303):
I think it could be made more clear that detrending was used only in the process of selecting strongly related climate variables, and not in the GLS models. Additionally, a comment on how the variability in the application of detrending method used here (thin plate splines) would impact climate variable selection would be warranted. There is a persistent question in dendrochronology about the amount of climate signal gets removed through the process of detrending.

*a bit of wording adjustment, and discuss detremding method (need help from dendro co-authors)*

**Regarding the detrending method, most detrending methods have been shown to retain the climate signal in many paleoclimate reconstructions. There are some ongoing nuanced discussions that are important for paleoclimate reconstructions such as retaining the medium frequency variance and examining the how segment length impacts that frequency, and about how changes near the end of the series (increasing or decreasing trend in growth) can impact the detrending and thus climate signal. However, for the purposes of this paper, the methods used here have been shown over and over again to retain the climate signal. We have added the following sentence to make this point: "These types of detrending methods are used to maximize the climate signal in tree-ring paleoclimate reconstructions and thus have been shown to retain the climate signal while removing the influence of other non-climatic drivers [@fritts_tree_1976; @cook_time_1985]."**

### 2021-07-28 doc/manuscript/ForestGEO_treerings_SI.Rmd
work on responses to reviewers
- DBH measurements
- detrending methods

Tree diameter at breast height (or height of coring), DBH, was reconstructed over the time span for which a core record existed.
DBH was reconstructed either from the outside in or from the inside out.
We gave precedence to the outside-in approach; that is, when a recent DBH measurement was available, DBH was reconstructed from the outside in. 
In cases where DBH was not available, but when we knew that the core hit pith or could reasonably estimate how far off it was based on the curvature of the rings [@applequist_simple_1958; @duncan_evaluation_1989], DBH was reconstructed from the inside out.

>>>>

Tree diameter at breast height (or height of coring), DBH, was reconstructed over the time span for which a core record existed.
DBH was reconstructed either from the outside in or from the inside out, depending on the availability of recent DBH measurements (see next paragraph).
We gave precedence to the outside-in approach; that is, when a recent DBH measurement was available, DBH was reconstructed from the outside in. 
In cases where DBH was not available, but when we knew that the core hit pith or could reasonably estimate how far off it was based on the curvature of the rings [@applequist_simple_1958; @duncan_evaluation_1989], DBH was reconstructed from the inside out.

DBH measurements were taken at the time of coring or -- for some of the trees cored within ForestGEO plots -- obtained from the most recent ForestGEO census prior to coring.
For these trees, we applied the outside-in approach to DBH reconstruction (detailed below).
For a minority of trees (n=50 at LDW), DBH measurements were not available, requiring use of inside-out approach to DBH reconstruction (detailed below).
